The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey

### 2020-02-26 RET_Hosts.tex

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[RET host galaxies in DES]{The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. Wiseman et al.]{
P. Wiseman$^1$\thanks{E-mail: p.s.wiseman@soton.ac.uk (PW)},
 M. Pursiainen$^1$,
 M. J. Childress$^1$,
 M. Smith$^1$,
 and Other Authors$^{1,2}$,
\newauthor
(The DES Collaboration)
\\
% List of institutions
$^{1}$School of Physics and Astronomy, University of Southampton, Southampton, SO17 1BJ, UK\\
$^{2}$Other Institutions\\
%$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/SFR_Mike.png}
\caption{The SFR of RET hosts, compared to CCSNe and the low-z SDSS sample.
\label{fig:sfms_sfr}}
\end{figure*}

@@

In order to subtract the stellar component of the host galaxy spectra, we use the Penalized PiXel-Fitting software (\texttt{pPXF}; \citealt{Cappellari2004,Cappellari2012,Cappellari2017}), using the MILES library of single stellar populations \citep{Vazdekis2010}. By subtracting the best-fitting stellar spectrum from the \texttt{pPXF} fit, we are left with a `gas' spectrum, comprising the emission lines. We fit the emission lines using \texttt{PySpecKit} \citep{Ginsburg2011}. In order to estimate the uncertainty on the emission line fluxes, we fit $10^4$ realisations of the line, each time adding a perturbation to the amplitude of the line by drawing from a Gaussian distribution centered on 0, and with a standard deviation equal to the root-mean-square (RMS) of the measured flux noise in that part of the spectrum. We take the mean and standard deviation of the resulting fits as our flux and its uncertainty, respectively.

@@

The most common method used to estimate the metallicity of galaxies is to use emission line ratios that have been calibrated using theoretical or empirical models in order to approximate the gas-phase oxygen abundance in the interstellar medium. Due to the low S/N of the spectra in this sample, we are constrained to a subset of these diagnostics by the availability of only a handful of the strongest emission lines, namely \halpha, \hbeta, \OII 3727, \OIII 4959 and 5007, \NII 6548 and 6583, and \SII 6717 and 6731. Furthermore, for each host galaxy only a subset of these lines are detected - for example, \halpha, \NII and \SII are redshifted out of the spectral coverage at $z>0.3$, leaving only the oxygen and \hbeta lines available, mandating the use of the R23 diagnostic. For hosts at $z<0.3$ we are able to use the \OIII /\NII (O3N2), \NII /\halpha (N2), and \SII /\NII (S2N2) line ratios.

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[RET host galaxies in DES]{The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. Wiseman et al.]{
P. Wiseman$^1$\thanks{E-mail: p.s.wiseman@soton.ac.uk (PW)},
 %M. Pursiainen$^1$,
 %M. J. Childress$^1$,
 %M. Smith$^1$,
 and Other Authors$^{1,2}$,
\newauthor
(The DES Collaboration)
\\
% List of institutions
$^{1}$School of Physics and Astronomy, University of Southampton, Southampton, SO17 1BJ, UK\\
$^{2}$Other Institutions\\
%$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/DES16C2ggt.png}
\caption{The spectrum of DES16C2ggt, decomposed into its constituent components according to the \texttt{pPXF} fit. 
\label{fig:host_spec}}
\end{figure*}

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/RET_BPT.png}
\caption{Baldwin-Phillips-Terlovich (BPT) diagram for RET hosts, showing that the emission lines are consistent with being generated by star-formation rather than AGN activity.
\label{fig:bpt}}
\end{figure}

@@

In order to subtract the stellar component of the host galaxy spectra, we use the Penalized PiXel-Fitting software (\texttt{pPXF}; \citealt{Cappellari2004,Cappellari2012,Cappellari2017}), using the MILES library of single stellar populations \citep{Vazdekis2010}. By subtracting the best-fitting composite stellar spectrum from the \texttt{pPXF} fit, we are left with a `gas' spectrum, comprising the emission lines. An example of this procedure is shown in Fig. \ref{fig:host_spec}. We fit the emission lines using \texttt{PySpecKit} \citep{Ginsburg2011}. In order to estimate the uncertainty on the emission line fluxes, we fit $10^4$ realisations of the line, each time adding a perturbation to the amplitude of the line by drawing from a Gaussian distribution centered on 0, and with a standard deviation equal to the root-mean-square (RMS) of the measured flux noise in a $100 \AA$ window around the central wavelength of the line. We take the mean and standard deviation of the resulting fits as our flux and its uncertainty, respectively.

@@

The most common method used to estimate the metallicity of galaxies is to use emission line ratios that have been calibrated using theoretical or empirical models in order to approximate the gas-phase oxygen abundance in the interstellar medium. Emission lines originate from regions of ionised gas, but there are a number of possible causes of this ionisation. Using the Baldwin-Phillips-Terlevich diagram (\ref{fig:bpt}; \citealt{Baldwin1981}), we demonstrate that the emission line ratios measured in RET hosts are consistent with ionisation caused by star-formation as opposed to AGN.

Due to the low S/N of the spectra in this sample, we are constrained to a subset of these diagnostics by the availability of only a handful of the strongest emission lines, namely \halpha, \hbeta, \OII 3727, \OIII 4959 and 5007, \NII 6548 and 6583, and \SII 6717 and 6731. Furthermore, for each host galaxy only a subset of these lines are detected - for example, \halpha, \NII and \SII are redshifted out of the spectral coverage at $z>0.3$, leaving only the oxygen and \hbeta lines available, mandating the use of the R23 diagnostic. For hosts at $z<0.3$ we are able to use the \OIII /\NII (O3N2), \NII /\halpha (N2), and \SII /\NII (S2N2) line ratios.
